---
title: "EDA"
output: pdf_document
date: '2022-04-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(R.matlab, stringr, tidyr, ggplot2, hexbin)

```

## questions
1. What does each file represent -- Zero_Maze has 13 sub folders with binned_zscore and binned_behavior. Is each a mouse?   
2. In the binned_behavior should each row have a 1 and a 0. Not 0,0. The mouse is either in the open arm or in the closed arm. There is no "other" option.   

## notes --
```binned_zscore.m``` = single cell recordings; col = each cell; row = each time  
```binned_behavior.m``` = animal behavior

Mouse Tasks:   
- Elevated Zero Maze; col1 = closed arm; col2 = open arm  
- Opposite Sex; col1 = interacting with male cup; col2 = interacting with female cup 
- Direct Interaction; col1 = interacting; col2 = non-social behavior/ rearing


# Zero Maze
```{r}
# read in our data
prefix <- "Data/Zero_Maze/"
folders <- list.files(prefix)

for (f in folders){
 #abbreviation for file name
 f_abbr <- str_sub(f, -3, -1)
 #subfolder
 sub_f <- list.files(paste0(prefix, f, "/Day_1/Trial_001_0"))
 
 #binned behavior
 assign(paste0("bb_",f_abbr), 
        readMat(paste0(prefix, f, "/Day_1/Trial_001_0/binned_behavior.mat")))
 
 #z score
 assign(paste0("bz_",f_abbr), 
        readMat(paste0(prefix, f, "/Day_1/Trial_001_0/binned_zscore.mat")))
  
}
```


 If I understand the data layout correctly, there are 32 neurons that are being tracked
 
```{r}

## Example with sub folder 251
x <- t(bb_251$binned.behavior)
z <- data.frame(bz_251$binned.zscore)


z$open <- x[,2]
z$closed <- x[,1]
z$time <- seq(1:nrow(z))

z_long <- z %>% pivot_longer(cols = X1:X32)

ggplot(data = z_long, mapping = aes(x = name, y = value, fill = as.factor(open)))+
  geom_violin(position = position_dodge())+
  labs(title = "Distributions of Neuron Z-Scores", fill = "Open?")+
    theme(legend.position = "bottom")


ggplot(data = z_long, mapping = aes(x = name, y = value))+
  geom_bin_2d(show.legend = FALSE, bins = 25, drop= FALSE)+
  scale_fill_viridis_c()+
  labs(title = "Heatmap of Activation: Open = 1, Closed = 0", x = "Neuron", y = "Z-score")+
  facet_wrap(vars(open), ncol = 1)+
  theme(legend.position = "bottom")

```

```{r}
x <- bb$binned.behavior
x <- bz$binned.zscore

```


